---
title: 'Session 1: R Basics'
author: "Callin"
date: "September 13, 2015"
output: html_document
---

# 1. Install R and R studio on your computer
          
*To download R, you go to this webpage: https://cran.r-project.org/ There, you can get both the Mac and Windows versions.
          
*You can download R studio here: https://www.rstudio.com/
          
*Once you have both installed, you should (hopefully) be able to click on RStudio, which will run R through a nice user interface.

### Check to see what version you have:
```{r}
version
```

# 2. Customize RStudio to make it how you like it
Preferences -> General -> save workspace to .RData upon exit -> Never

Color

Tabs, etc

# 3. Use R as a calculator
```{r}
3 + 5
exp(44)
log(90)
1:5 + 3
3 > 5
```

# 4. Write an R script
File -> New File -> R Script

# 5. Comment your code

"#"" is the comment character

Comments won't be evaluated
```{r}
#-#-#-#-#-#-#-#-#-#-#-#-#-#-#-#-#-#-#-#-#-#-#-#-#-#-#-#-#-#-#-#-#-
# 4 + 3
4 + 3
```

# 6. Assign values to variables
```{r}
var1 <- "Hi"
var2 <- 4
var3 <- "Work that up-do!"
var4 <- matrix(data = sample(x = 1:100), ncol = 10)

var1
var2
var3
var4

# matrix multiplication
var4 %*% matrix(1:10, ncol = 1)

# matrix inverting
solve(var4) 

# matrix transpose
t(var4)

```

# 7. Use different data types – factor, character, numeric, logical
```{r}
# factors
factor1 <- factor("a")
class(factor1)

# Character / Strings
## often, a sequence of characters is called a string, but R doesn't really differentiate
ch1 <- "Hi" # quotes
class(ch1)

ch2 <- "4" # quotes mean characters
class(ch2)

# Numeric
num1 <- 4
class(num1)

num2 <- as.numeric(ch2)
class(num2)

# Logical
logic1 <- TRUE
class(logic1)

logic2 <- FALSE
class(logic2)

logic3 <- as.logical(1) # can convert 1 and 0 to TRUE / FALSE
logic3

# NULL
n1 <- NULL
class(n1)

# Side Note:
# factors can be confusing
factor1 <- factor(x = 0)
as.numeric(factor1) # weird...

# here's why:
f2 <- factor(c("A", "B", "C", "D", "A", "D"))
f2
as.numeric(f2) # note that R indexes, starting with 1, instead of 0

# here's how to convert factors to numbers exactly
as.numeric(as.character(factor1)) # better
```


# 8. Use different data structures – vector, matrix, array, list, data frame

There are some other structures, like tables, but we'll skip that for now.

### Vectors
```{r}
# Vectors are one dimensional arrays with one type of data (i.e. can't mix logical and character)
# In math, we often write them like this: < 1, 3, 4 >
# In R, we "concatenate" individual elements together to make vectors

vec1 <- c(1, 3, 4)
vec1

# get info about our variable
str(vec1)
class(vec1)
mode(vec1)
length(vec1)
summary(vec1)

is.vector(vec1)
is.atomic(vec1)

## vectors can be different data types (but not factors, since they have levels)
# logical vector
lv1 <- c(TRUE, TRUE, FALSE, FALSE, TRUE)
class(lv1)
str(lv1)
is.vector(lv1)

# Character vector
cv1 <- c("Hi", "Everyone!")
cv1
class(cv1)
str(cv1)
length(cv1)
is.vector(lv1)

# can't have factor vectors
fv1 <- as.factor(cv1)
fv1
is.vector(fv1)

```

### Matrix
```{r}
# matrices are for single data types
m1 <- matrix(data = sample(1:36), nrow = 6)
m1 # numeric matrix
mode(m1)
class(m1)

# character matrix
chars <- sample(letters, 36, replace = TRUE)
chars
matrix(chars, nrow = 6)

# logical matrix
logic4 <- c(TRUE, FALSE, TRUE, TRUE)
matrix(logic4, nrow = 2)
matrix(logic4, nrow = 2, byrow = TRUE) # input by rows, rather than columns

# matrix() converts factors to characters
factChars <- as.factor(chars)
factChars
factM1<- matrix(factChars, nrow = 6)
factM1
mode(factM1)

```

### Arrays are like stacks of matrices
```{r}
m2 <- matrix(rep(2, 36), nrow = 6)

array(data = c(m1, m2), dim = c(6,6,2))
```


### Lists can contain anything!
navigate lists wiht double brackets -- [[ ... ]]
```{r}
l1 <- list(n1, logic2, m2, factM1, factChars, num2)
l1

l1[[1]]
l1[[4]]

```

### Data frames are lists of vectors (or factors) that are all the same length

Data frames are usually organized by their columns
Data frames can have factors, numbers, characters, etc.
```{r}
df1 <- as.data.frame(m1)
df1

# a little of what is to come
# special data frame navigation with "$"
df1$V1
mode(df1$V1)
class(df1$V1)

# proof that data frames are lists -- you can use [[ ... ]]
df1[[1]]
df1[["V1"]] # can use column names to navigat them
df1["V1"] # using single brackets might not get you what you want
mode(df1["V1"]) # not a vector
df1[, "V1"] # if you put a comma in it works?  More on this later!
mode(df1[["V1"]])
```


# 9. Use pre-written functions to find the characteristics of variables (length, names, etc.)
```{r}
# let's start with a more interesting data frame
df1$V1 <- as.factor(df1$V1) # convert to factor

# add a column to df
df1$V7 <- sample(c(TRUE, FALSE), replace = TRUE, size = nrow(df1))

head(df1)
summary(df1)
str(df1)
class(df1)
mode(df1)

# get number of columns
length(df1) 
ncol(df1)

# get number or rows
nrow(df1) 
length(df1$V1)

# get names of columns
colnames(df1)
names(df1)

# change column names
colnames(df1) <- c("One", "Two", "Cat", "Foo", "Dog", "Bar", "Moo")
head(df1)
```

# 10.   Import your data into R
You can click "Import Dataset", or write the code yourself
```{r}
## Here is some of my unpublished data -- if you publish it, please include me as an author!
## but seriously, please don't share.
# flowerFile <- file.choose() 

flowerFile <- "/Users/callinswitzer/Dropbox/Harvard/RByCallin/Session1/FlowerPollenExperiment-ExampleRClass.csv"

flr <- read.csv(flowerFile)
head(flr)

## you can also import directly from a web URL
# this is a dataset about rats from crawley book
URL <- "http://nature.berkeley.edu/~casterln/crawley/rats.txt"
rats <- read.csv(URL)
head(rats) # not quite right

rats <- read.csv(URL, sep = "\t")
head(rats)
```

# 11.   Set and get your working directory
Session -> Set Working Directory -> Choose Directory
```{r}
getwd() # get working directory

dir() # print what's in your directory

setwd("Session1") # can set a new director by typing a folder in your current wd.
getwd()

setwd("..") # goes "up" in your directory
getwd()

```

# 12. Navigate your data, using “square bracket” notation and “$” notation.
```{r}
head(flr)

head(flr["frameRate"]) # this returns a data frame, which we often don't want
flr[["frameRate"]][1:10] # this returns a vector, first 10 items


# if you want a new data frame with a subset of columns
smDF <- flr[c("frameRate", "flowerNum")]
head(smDF)
# another possible way
smDF <- flr[, c("frameRate", "flowerNum")] # [rows, columns]
head(smDF)
# another way
smDF <- flr[, c(19,1)] # getting all rows, columns 19 and 1
head(smDF)

# if you just want a single column from your data frame
shutter <- flr$shutter
shutter[1:10] # gives first 10 items from the "shutter" variable
```

# 13. Quickly visualize your data (using car::scatterplotMatrix() and plot())
```{r}
plot(x = flr$humidity, y=flr$temp)

colnames(flr)
#install.packages("car")
car::scatterplotMatrix(flr[c("humidity", "temp", "slope", "int")])
```

# 14. Manipulate your data (sub setting with logical statements, cleaning data, making calculations with a data frame)
```{r}
# select rows, all columns
FewRows <- flr[10:20, ]
FewRows

# select rows where daysOpen == 1, all columns
d1 <- flr[flr$daysOpen == 1, ]
head(d1)

# select rows where daysOpen does not = 1, and distance == 1
d1d1 <- flr[flr$daysOpen != 1 & flr$distance == 1, ]
head(d1d1)


# cleaning data
hist(flr$slope); rug(flr$slope)

# say we want to drop the points where slope > 0.12
sub1 <- flr[flr$slope < 0.12, ]
hist(sub1$slope); rug(sub1$slope)

# say we want to calculate a new measurement that is a combination of slope + int
flr$NewStat <- flr$slope + flr$int
head(flr$NewStat)

# or this will work
flr$NewStat1 <- with(flr, slope + int)
head(flr$NewStat1)

# say we want to change one of the variable types in our data frame
class(flr$daysOpen)
flr$daysOpen <- as.factor(flr$daysOpen)
class(flr$daysOpen)
# or this will work
class(flr$flowerNum)
flr <- within(flr, flowerNum <- as.factor(flowerNum))
class(flr$flowerNum)


# calculate averages by individual
# say we want average glycogen for each individual
summary(rats)
tapply(X = rats$Glycogen, FUN = mean, INDEX = rats$Rat)

# calcuate mean by treatment and individual
tapply(X = c(rats$Glycogen), FUN = mean, INDEX = interaction(rats$Rat, rats$Treatment))


tab1 <- tapply(X = flr$slope, INDEX = flr$trt, FUN = mean)
class(tab1)
mode(tab1)

# tabulate your data (get counts)
xt1 <- xtabs(formula = ~trt, data = flr)
xt1

# convert to data frame for easy saving
tab1 <- as.data.frame(tab1)


```

# 15. Export data
```{r}
# Write text files
# can also use write.csv()
write.table(x = tab1, file = "Tab1.csv", sep = ",",  col.names = FALSE)

# saving plots
pdf(file = "plot1.pdf", width = 5, height = 4)
plot(x = flr$humidity, y=flr$temp)
dev.off()

# or just export with the "Export" button 
```



