---
title: "Survey Analysis"
author: "Callin Switzer"
date: "September 11, 2015"
output: html_document
---

#### Read in data
```{r}
#direct <- file.choose() # handy way to change directory
direct <- "/Users/callinswitzer/Dropbox/Harvard/RByCallin/R Course for OEB (Responses) - Form Responses 1.csv"
surv <- read.csv(direct, stringsAsFactors = FALSE)
```

#### Find how many people responded
```{r}
nrow(surv)
```

Great! This shows that `r nrow(surv)` people responded!

#### Find what time works best for folks
```{r}
# separate the possible times into a single vector
TIMES <- unlist(strsplit(surv$What.times.would.work.for.you.to.participate., split = ", "))

# make it a data frame, so we can plot it with ggplot
TimesDF <- data.frame(TimePref = TIMES)
library(ggplot2) # load library

# set the levels, to reorder the bars, tallest being on the left
TimesDF$TimePref <- factor(TimesDF$TimePref, levels = names(sort(table(TimesDF$TimePref), 
                               decreasing=TRUE)))

# plot it
aa <- ggplot(TimesDF, aes(x = TimePref)) + 
     geom_bar() + #barplot
     theme_bw() + # black and white theme
     theme(axis.text.x=element_text(angle=45, hjust = 1)) # rotate text on x axis
aa
```
This plot is okay, but I'll fix it up a bit.


#### define a custom theme for plotting
```{r}
library(grid) # for the unit() function
my.theme <- theme_bw() + 
               theme(plot.background = element_blank(),
                    panel.grid.major = element_blank(),
                    panel.grid.minor = element_blank(),
                    panel.border = element_blank(),
                    axis.line = element_blank(),
                    axis.text.y = element_text(size=10),
                    axis.ticks.length=unit(-0.15, "cm"),
                    axis.ticks.margin=unit(0.4, "cm"), 
                    legend.title=element_blank(), 
                    legend.position = "none", 
                    legend.justification = c(1, 1),
                    axis.ticks.y = element_line(),
                    axis.text.x=element_text(angle=45, hjust = 1), 
                    axis.ticks.x = element_blank(), 
                    plot.margin = unit(c(2,1,1,1), "cm")) 
```


#### make some plots
```{r, tidy=TRUE}
# make a plot with weird colors
plot1 <- ggplot(TimesDF, aes(x = TimePref)) + 
     geom_bar(alpha = 0.8)      
     
p1 <- plot1 +  my.theme + labs(x = "Time Preferences", y = "Frequency")
p1

# save plot
ggsave("timePrefs.png", p1, width = 10, height = 8)
```


#### find who can and can't come on 2nd most popular day (Tuesday)
```{r}
# scrample names, for PRIVACY!
NameScramble <- function(o) paste(sample(unlist(strsplit(surv$Name., split = "")), size = rpois(1, lambda = 14)), collapse = "")
surv$Name. <- replicate(nrow(surv), expr = NameScramble())

WhoCanCome <- grepl(pattern = levels(TimesDF$TimePref)[2], x = surv$What.times.would.work.for.you.to.participate.)

# here are the people who can come
surv$Name.[ WhoCanCome]

# find who cannot come
surv$Name.[!WhoCanCome]
```


# visualize people's experience with R
```{r}
# visualize experience
lvs <- levels(factor(surv$How.would.you.rate.your.experience.with.coding.in.R.))
surv$How.would.you.rate.your.experience.with.coding.in.R. <- factor(surv$How.would.you.rate.your.experience.with.coding.in.R., levels = lvs[c(3,1,4,2)])


expPlot <- ggplot(surv, aes(x = How.would.you.rate.your.experience.with.coding.in.R.)) + 
     geom_bar() + 
     my.theme + 
     scale_y_continuous(expand = c(0,0))
expPlot

# save the plot
ggsave("ExpPlot.png", expPlot, width = 10, height = 8)
```


#### plot most commonly used words that excite people about R
```{r}
excite <- surv$What.are.you.most.excited.to.learn..using.R.
#split by space or return
wordsUsed <- (unlist(strsplit(excite, split = " |\n")))


# remove punctuation 
wdsCln <- tolower(gsub("[[:punct:]]", replacement = "", x = wordsUsed))

# remove blanks, and useless words. 
wdsCln <- data.frame(wds = wdsCln[wdsCln != ""])
articles <- c("to", "and", "a", "of", "for", "the", "in", "with", "be", "also", "some", "as", "so", "at", "although", "this", "all", 'it', "but", "is", "i", "very", "that", "would", "am", "really", "about")
wdc <- wdsCln$wds[!(wdsCln$wds %in% articles)]

# get top 25 words
wrdTab <- sort(table(wdc), decreasing = TRUE)[1:25]

# make a data frame
wrdDF <- data.frame(BuzzWords = wdc[wdc %in% names(wrdTab)])

# relevel the factor so that the most popular graphs to the left side
wrdDF <- within(wrdDF, 
                BuzzWords <- factor(BuzzWords, levels = names(wrdTab)))

# plot it
LearnPlot <- ggplot(wrdDF, aes(x=BuzzWords)) + 
     geom_bar(alpha = 0.9) + 
     my.theme + 
     theme(axis.ticks.x = element_blank()) + 
     scale_y_continuous(expand = c(0,0)) + 
     labs(x = "Words used to describe interest", y = "Frequency")
     
LearnPlot

# save the plot 
ggsave(filename = "LearnPlot.png", plot = LearnPlot, width = 10, height = 8)
```

#### Make a map to show people where to go!
```{r}
# HUH coordinates
coords <- c(42.379007, -71.114572)
# loading the required packages
library(ggmap)

# make a data frame
df <- data.frame(lat = coords[1], lon = coords[2])

# getting the map
mapHUH <- get_map(location = c(lon = mean(df$lon), lat = mean(df$lat)), zoom = 18,
                      maptype = "hybrid", scale = 2)

# plotting the map with some points on it
m1 <- ggmap(mapHUH) +
     geom_point(data = df, aes(x = lon, y = lat),  shape = 18 , size = 10, color = "red") +
     guides(fill=FALSE, alpha=FALSE, size=FALSE)
m1

# save
ggsave("MapToHUH.png", m1, width = 5, height = 5)
```


```{r, eval = F, echo = F}
# print out email addresses for interested parties
paste(surv$Preferred.Email.Adress., collapse = ", ")
is.character("bob")
```